NYC Spatial Models - Comparison of Spatial Models With and Without Twitter Data

### 2018-11-02 NYC_Twitter_SDM_models - Copy.Rmd
Look for a better transformation for tweets variable

@@

Calculate the logsum given a tracts and parks dataset. `tracts$park_ls1` uses
only size and distance parameters. `tracts$park_ls2` uses size, distance and
tweet parameters. `tracts$park_ls4` and `tracts$park_ls5` use a distance
parameter of -15.
```{r accessibility}
# calculate centroids of tracts and parks
tract_centroids <- gCentroid(tracts, byid = TRUE) 
park_centroids <- gCentroid(parks, byid = TRUE)

# distances from all tracts to all zones
distances <- gDistance(park_centroids, tract_centroids, byid = TRUE) / 5280

tracts$park_ls1 <- calculate_park_logsums(tracts, parks, distances,
                                          parks$Park_Acres, betas = c(.00001, -5))

tracts$park_ls2 <- calculate_park_logsums(tracts, parks, distances, 
                                          parks$Park_Acres, parks$TWEET_COUNT, 
                                          betas = c(.00001, -5, .001))

@@

tracts$park_ls5 <- calculate_park_logsums(tracts, parks, 
                                          parks$Park_Acres, parks$TWEET_COUNT, 
                                          betas = c(.00001, -15, .001))
```

Map the newly created Park Access variables. Here is a map of accessibility using only the size and distance parameters that are most likely when using obesity as the dependent variable.
```{r accessibility_leaflet_1, echo=FALSE}
# Create a continuous palette function
pal <- colorQuantile(
  palette = "Greens",
  domain = c(0, tracts$park_ls1)
)

pop_content <- function(shp){
  paste0(
    "Tract: ", shp$NAME, "<br/>",
    "% Park: ", round(shp$Park_Percent, 2), "<br/>",
    "Park Access: ", round(shp$park_ls1, 2), "<br/>"
  )
}

leaflet(tracts %>% spTransform(., CRS("+init=epsg:4326"))) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(weight = 1, popup = ~pop_content(tracts),
              color = ~pal(park_ls1))
```

And here is a map of accessibility using the size, distance and tweet parameters (using obesity as the DV). 
```{r accessibility_leaflet_2, echo=FALSE}
# Create a continuous palette function
pal <- colorQuantile(
  palette = "Greens",
  domain = c(0, tracts$park_ls2)
)

pop_content <- function(shp){
  paste0(
    "Tract: ", shp$NAME, "<br/>",
    "% Park: ", round(shp$Park_Percent, 2), "<br/>",
    "Park Access: ", round(shp$park_ls2, 2), "<br/>"
  )
}

leaflet(tracts %>% spTransform(., CRS("+init=epsg:4326"))) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(weight = 1, popup = ~pop_content(tracts),
              color = ~pal(park_ls2))
```

Here is a map of accessibility using the betas that produce the most likely model when using physical activity as the dependent variable. Only size and distance parameters are used.
```{r accessibility_leaflet_3, echo=FALSE}
# Create a continuous palette function
pal <- colorQuantile(
  palette = "Greens",
  domain = c(0, tracts$park_ls4)
)

pop_content <- function(shp){
  paste0(
    "Tract: ", shp$NAME, "<br/>",
    "% Park: ", round(shp$Park_Percent, 2), "<br/>",
    "Park Access: ", round(shp$park_ls4, 2), "<br/>"
  )
}

leaflet(tracts %>% spTransform(., CRS("+init=epsg:4326"))) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(weight = 1, popup = ~pop_content(tracts),
              color = ~pal(park_ls4))
```

And here is a map of accessibility using size, distance and tweet parameters (with the combination of betas that is most likely for physical activity as the DV).
```{r accessibility_leaflet_4, echo=FALSE}
# Create a continuous palette function
pal <- colorQuantile(
  palette = "Greens",
  domain = c(0, tracts$park_ls5)
)

pop_content <- function(shp){
  paste0(
    "Tract: ", shp$NAME, "<br/>",
    "% Park: ", round(shp$Park_Percent, 2), "<br/>",
    "Park Access: ", round(shp$park_ls5, 2), "<br/>"
  )
}

leaflet(tracts %>% spTransform(., CRS("+init=epsg:4326"))) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(weight = 1, popup = ~pop_content(tracts),
              color = ~pal(park_ls5))
```

Create regression formulas:
```{r obese_lm, echo=FALSE, message=FALSE, include=FALSE}
#base_formula <- formula(~log(Pop_Density) + FulltimeWork + 
    #CollegeDeg + Single_Percent + PctWhite + PctBlack + PctAsian + PctOther + PctHispanic)
base_formula_age <- formula(~log(Pop_Density) + FulltimeWork + 
    CollegeDeg + Single_Percent + Pct0to17 + Pct18to29 + Pct30to64 + Pct65plus + PctWhite + PctBlack + PctAsian + PctOther + PctHispanic)

#parks_formula_1 <- update(base_formula, . ~ . + log(park_ls1 + 0.1))
parks_formula_1a <- update(base_formula_age, . ~ . + log(park_ls1 + 0.1))
#parks_formula_2 <- update(base_formula, . ~ . + log(park_ls2 + 0.1))
parks_formula_2a <- update(base_formula_age, . ~ . + log(park_ls2 + 0.1))
#parks_formula_3 <- update(base_formula, . ~ . + log(park_ls4 + 0.01))
parks_formula_3a <- update(base_formula_age, . ~ . + log(park_ls4 + 5))
#parks_formula_4 <- update(base_formula, . ~ . + log(park_ls5 + 0.1))
parks_formula_4a <- update(base_formula_age, . ~ . + log(park_ls5 + 5))

>>>>

The Tweets count variable has a high degree of zeros (virtually all parks have
no tweets originating there). We should identify if there is a transformation
that will make the distribution more normal (log is typical), but that will allow
for zero variables as $log(0)$ is undefined.

```{r tweettransformation}
parks_sf %>%
  st_set_geometry(NULL) %>%
  tbl_df() %>%
  select(raw = TWEET_COUNT) %>%
  mutate(
    `0.5` = yeo.johnson(raw, 0.5),
    `0` = yeo.johnson(raw, 0),
    log = log(raw + 1)
  ) %>%
  sample_n(200) %>%
  gather(lambda, x, -raw) %>%
  ggplot(aes(x = raw, y = x, color = lambda)) + 
  geom_point()
```

@@

```{r distancematrix}
# calculate centroids of tracts and parks
tract_centroids <- gCentroid(tracts, byid = TRUE) 
park_centroids <- gCentroid(parks, byid = TRUE)

# distances from all tracts to all zones
distances <- gDistance(park_centroids, tract_centroids, byid = TRUE) / 5280
```

Calculate the logsum given a tracts and parks dataset. `tracts$park_ls1` uses
only size and distance parameters. `tracts$park_ls2` uses size, distance and
tweet parameters. `tracts$park_ls4` and `tracts$park_ls5` use a distance
parameter of -15.

We need to firm up these calculations. Are we using log of acres as well?

```{r logsums}
tracts$park_ls1 <- calculate_park_logsums(tracts, parks, distances,
                                          parks$Park_Acres, betas = c(.00001, -5))

tracts$park_ls2 <- calculate_park_logsums(tracts, parks, distances,
                                          parks$Park_Acres,
                                          yeo.johnson(parks$TWEET_COUNT, 0),
                                          betas = c(.00001, -5, 2))

@@

tracts$park_ls5 <- calculate_park_logsums(tracts, parks, distances,
                                          parks$Park_Acres,
                                          yeo.johnson(parks$TWEET_COUNT, 0),
                                          betas = c(.00001, -15, 2))
```

I'm interested in looking at the distribution of the log-sum statistics so we can
know what if anything the different variables are doing, also to make sure that
there is no missingness in the data.

  - `ls1` and `ls2` have virtually the same distribution, along with `ls4` and
  `ls5`. This suggests that tweets have virtually no effect on the rankings, at
  least in this iteration.

```{r distribution}
logsums <- tracts@data %>%
  select(park_ls1:park_ls6) %>%
  gather(logsum_type, value)

# check missingness
table(is.na(logsums$value))

# histogram
ggplot(logsums, aes(x = value, fill = logsum_type)) + geom_histogram()
```

Map the newly created Park Access variables. Here is a map of accessibility using only the size and distance parameters that are most likely when using obesity as the dependent variable.
```{r accessibility_leaflet_1, echo=FALSE}
# Create a continuous palette function
pal1 <- colorQuantile(palette = "Greens", domain = tracts$park_ls1)
pal2 <- colorQuantile(palette = "Greens", domain = tracts$park_ls2)
pal4 <- colorQuantile(palette = "Greens", domain = tracts$park_ls4)
pal5 <- colorQuantile(palette = "Greens", domain = tracts$park_ls5)

leaflet(tracts %>% spTransform(., CRS("+init=epsg:4326"))) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(group = "Long distance, no tweets", weight = 1, color = ~pal1(park_ls1)) %>%
  addPolygons(group = "Long distance, tweets", weight = 1, color = ~pal2(park_ls2)) %>%
  addPolygons(group = "Short distance, no tweets", weight = 1, color = ~pal4(park_ls4)) %>%
  addPolygons(group = "Short distance, tweets", weight = 1, color = ~pal5(park_ls5)) %>%
  addLayersControl(baseGroups =  c("Short distance, no tweets", 
                                   "Short distance, tweets", 
                                   "Long distance, no tweets", 
                                   "Long distance, tweets"))
```

A
Create regression formulas:
```{r obese_lm, echo=FALSE, message=FALSE, include=FALSE}
#base_formula <- formula(~log(Pop_Density) + FulltimeWork + 
    #CollegeDeg + Single_Percent + PctWhite + PctBlack + PctAsian + PctOther + PctHispanic)
base_formula_age <- formula(~log(Pop_Density) + FulltimeWork + 
    CollegeDeg + Single_Percent + Pct0to17 + Pct18to29 + Pct30to64 + Pct65plus + PctWhite + PctBlack + PctAsian + PctOther + PctHispanic)

parks_formula_1a <- update(base_formula_age, . ~ . + park_ls1 )
parks_formula_2a <- update(base_formula_age, . ~ . + park_ls2 )
parks_formula_3a <- update(base_formula_age, . ~ . + park_ls4 )
parks_formula_4a <- update(base_formula_age, . ~ . + park_ls5 )
